---
title: "Person-Centered Measurement"
author: Uses of Assessment Data to Inform Services for Individuals with Intellectual and Developmental Disabilities
output:
  html_document:
    self_contained: no
    theme: flatly
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
always_allow_html: yes
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse); library(plotly); library(readxl); library(stringr); library(forcats); library(magrittr)

# Read in focus group summary data
sis_focus <- read_excel("focus_group_quotes.xlsx")

# Read in survey data
sis_survey <- read.csv("survey_results/CSV/Sheet_1.csv", na.strings = c("NA",""))

# Clean survey and rm unnecessary fields

sis_survey %<>% 
  # Remove columns with all NAs
  select_if(colSums(!is.na(.)) > 0) %>%
  mutate(RespondentID = as.factor(RespondentID),
         CollectorID = as.factor(CollectorID)) %>%
  # Remove test collector results
  filter(CollectorID != "152740888") %>%
  # Rename fields for readability
  rename(role = Which.statement.best.describes.your.involvement.in.Intellectual.and.Developmental.Disabilities..IDD..services.in.Michigan.,
         use_access = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Make.sure.people.receive.the.services.they.need,
         use_networkdev = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Understand.population.level.needs.for.development.of.provider.network.capacity,
         use_networkadeq = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Evaluate.network.adequacy.for.people.with.IDD,
         use_geography = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Identify.geographic.areas.to.increase.service.availability.or.build.new.programs,
         use_popdiff = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Understand.differences.in.needs.across.the.population,
         use_iposneeds = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Recommend.specific.needs.for.inclusion.in.an.individual.plan.of.service..IPOS.,
         use_iposgoals = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Inform.the.person.centered.planning.process.by.identifying.potential.goals.and.objectives,
         use_refer = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Identify.potential.service.referrals.for.individuals,
         use_safetyplan = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Help.guide.safety.planning,
         use_predictcost = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Predict.costs,
         use_newservice = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Forecast.a.need.for.services.which.may.not.currently.be.available,
         use_naturalsupport = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Identify.people.who.may.be.at.risk.if.they.lose.a.large.portion.of.natural.supports..e.g..aging.parents.,
         use_consistentauth = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Identify.cases.where.services.being.authorized.or.provided.are.inconsistent.with.a.person.s.needs,
         use_itemintent = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Ensure.that.assessors.are.consistently.understanding.item.intent,
         use_timely = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Ensure.SISÂ..assessments.are.being.completed.in.a.timely.manner,
         use_valid = How.could.the.data.from.the.SISÂ..assessment.help.you.to.perform.your.role....Review.validity.and.reliability.of.SISÂ..assessments,
         review_IRQR = Please.rate.the.frequency.at.which.you.review.the.following.data....I.review.SISÂ..results..assessor.ratings.compared.to.regional.ratings..as.a.part.of.the.Interviewer.Reliability.and.Qualification.Review..IRQR..process.,
         review_drift = Please.rate.the.frequency.at.which.you.review.the.following.data....I.review.SISÂ..results..assessor.ratings.compared.to.regional.ratings..as.a.part.of.the.Periodic.Drift.Review.process.,
         review_compliance = Please.rate.the.frequency.at.which.you.review.the.following.data....I.audit.local.SISÂ..data.for.compliance.and.procedural.drift.markers.,
         family_receives = The.individuals.families.I.serve.receive.the.SISÂ..Family.Friendly.report.by.,
         review_glance = I.review.the.SISÂ..assessment.report.by....Glancing.over.it,
         review_importtofor = I.review.the.SISÂ..assessment.report.by....Briefly.reviewing.the..important.to..and..important.for..sections,
         review_readall = I.review.the.SISÂ..assessment.report.by....I.read.the.entire.report,
         review_makenotes = I.review.the.SISÂ..assessment.report.by....Making.notes.of.items.to.include.in.person.centered.planning.process,
         review_compareyears = I.review.the.SISÂ..assessment.report.by....Comparing.needs.to.goals.from.last.year.s.plan.and.make.notes.on.any.gaps.or.changes,
         review_none = I.review.the.SISÂ..assessment.report.by....I.don.t.review.the.SIS.assessment.report,
         review_other = I.review.the.SISÂ..assessment.report.by....Other..please.specify.,
         review_otherdesc = I.review.the.SISÂ..assessment.report.by....Other..please.specify..1,
         sc_goals = Please.identify.the.item.which.best.reflects.your.use.of.information.from.the.SISÂ.....I.use.the.identified.needs.in.the.SISÂ..assessment.report.to.identify.potential.goals.and.objectives,
         sc_services = Please.identify.the.item.which.best.reflects.your.use.of.information.from.the.SISÂ.....I.use.the.identified.needs.in.the.SISÂ..assessment.report.to.identify.potential.service.referrals,
         sc_pcp = Please.identify.the.item.which.best.reflects.your.use.of.information.from.the.SISÂ.....I.embed.the.review.of.SISÂ..assessment.results.into.my.preparation.for.or.facilitation.of.person.centered.planning.meetings,
         sc_amtscpdur = Please.identify.the.item.which.best.reflects.your.use.of.information.from.the.SISÂ.....I.use.the.information.found.in.the.SISÂ..assessment.report.to.inform.the.amount..scope..and.duration.of.services.requested.in.the.Individual.Plan.of.Service..IPOS.,
         sc_whynot = Please.identify.the.item.which.best.reflects.your.use.of.information.from.the.SISÂ.....If.you.answered.Never.or.Occasionally..please.explain.why.,
         time_review = I.spend.the.following.amount.of.time..average.per.person.family...........reviewing.the.SISÂ..assessment.report,
         time_explain = I.spend.the.following.amount.of.time..average.per.person.family...........explaining.the.SISÂ..results,
         time_whynot = I.spend.the.following.amount.of.time..average.per.person.family.......If.you.selected.None..please.explain.why.,
         qual_explain = Please.indicate.your.level.of.agreement.with.the.following.statements....I.am.confident.I.provide.a.helpful.and.thorough.explanation.of.the.SISÂ..assessment.results.to.persons.served.family.members.,
         qual_pcptrain = Please.indicate.your.level.of.agreement.with.the.following.statements....I.have.received.adequate.training.on.how.to.understand.and.discuss.SISÂ..assessment.results.and.how.SISÂ..assessment.information.should.be.used.in.the.context.of.person.centered.planning.,
         qual_whynot = Please.indicate.your.level.of.agreement.with.the.following.statements....If.you.selected.Strongly.Disagree.or.Disagree..please.explain.why.,
         deferneed = I.use.SISÂ..assessment.data.to.ensure.any.identified.need.that.is.not.addressed.in.the.individual.plan.of.service..IPOS..is.deferred.by.,
         deferneed_whynot = I.use.SISÂ..assessment.data.to.ensure.any.identified.need.that.is.not.addressed.in.the.individual.plan.of.service..IPOS..is.deferred.by....If.you.selected..I.don.t.use.SIS.assessment.data.for.this...please.explain.why.,
         integrate_pcp_open = What.is.the.most.effective.and.efficient.way.to.incorporate.review.of.SISÂ..assessment.results.into.clinical.processes.and.person.centered.planning....Open.Ended.Response,
         cls_role = The.SISÂ..assessment.results.help.me.understand.my.role.as.a.CLS.worker.and.how.I.can.better.support.the.individual.with.whom.I.am.working.,
         cls_whynot = The.SISÂ..assessment.results.help.me.understand.my.role.as.a.CLS.worker.and.how.I.can.better.support.the.individual.with.whom.I.am.working....If.you.selected.Strongly.Disagree.or.Disagree..please.explain.why.,
         um_locreview = I.review.SISÂ..assessment.results.to.inform.my.utilization.management.decisions....When.completing.a.UM.record.review.to.determine.if.appropriate.level.of.care.is.being.provided,
         hsw_identify = My.region.has.a.process.for.using.the.SISÂ..data.to.identify.individuals.who.may.benefit.from.HSW.services.,
         super_review = Please.rate.the.following.based.on.your.experience.within.the.programs.you.oversee....Timely.completion.of.SISÂ..assessments.is.reviewed.in.supervision.with.Supports.Coordinators.Case.Managers.on.a.regular.basis.,
         super_explain = Please.rate.the.following.based.on.your.experience.within.the.programs.you.oversee....Supports.Coordinators.Case.Managers.have.received.adequate.training.on.how.to.understand.SISÂ..assessment.results.and.explain.these.to.consumers.family.members.,
         super_pcptrain = Please.rate.the.following.based.on.your.experience.within.the.programs.you.oversee....Supports.Coordinators.Case.Managers.receive.adequate.training.on.how.to.use.SISÂ..assessment.results.to.inform.the.person.centered.planning.process.,
         super_whynot = Please.rate.the.following.based.on.your.experience.within.the.programs.you.oversee....If.you.selected.Strongly.Disagree.or.Disagree..please.explain.why.,
         direct_timely = Supervisors.within.the.programs.I.oversee.utilize.a.process.for.monitoring.the.timeliness.of.SISÂ..family.friendly.reports.being.provided.to.families.and.SISÂ..assessment.reports.being.provided.to.Support.Coordinators.Case.Managers.,
         family_friendly = The.family.friendly.SISÂ..summary.reports.are.helpful.and.inform.the.person.centered.plan.,
         participated = Did.you.participate.in.a.SISÂ..assessment.within.the.past.six.months.,
         participant_purpose = Please.complete.the.following.items.as.they.apply.to.the.SISÂ..assessment.in.which.you.participated....The.purpose.of.the.assessment.was.explained.to.me.clearly.,
         participant_answers = Please.complete.the.following.items.as.they.apply.to.the.SISÂ..assessment.in.which.you.participated....The.assessor.was.friendly.and.helpful.in.answering.my.questions.,
         participant_respect = Please.complete.the.following.items.as.they.apply.to.the.SISÂ..assessment.in.which.you.participated....I.was.treated.with.dignity.and.respect.throughout.the.process.,
         participant_understand = Please.complete.the.following.items.as.they.apply.to.the.SISÂ..assessment.in.which.you.participated....The.assessment.helped.me.to.better.understand.the.needs.of.the.individual.being.assessed.,
         participant_results = Please.complete.the.following.items.as.they.apply.to.the.SISÂ..assessment.in.which.you.participated....I.understood.the.results.of.the.assessment.,
         participant_received = I.received.the.results.of.the.assessment.in.a.report.,
         overall = IDD.Services.within.Michigan.are.effectively.using.SISÂ..assessment.data.to.inform.and.improve.care.,
         overall_whynot = IDD.Services.within.Michigan.are.effectively.using.SISÂ..assessment.data.to.inform.and.improve.care....If.you.selected.Strongly.Disagree.or.Disagree..please.explain.why.) %>%
  # Recode Roles
  mutate(role = recode(role,
                       `I am a Supports Coordinator/Case Manager serving individuals with IDD.` = "Supports Coordinator",
                       `I am a Supervisor/Director of programs for individuals with IDD.` = "Program Director",
                       `I am a supervisor of SISÂ® Assessor(s).` = "Assessor Supervisor",
                       `I am a regional SISÂ® Quality Lead and/or a SISÂ® Assessor.` = "SIS Assessor or Quality Lead",
                       `I am an advocate for individuals with IDD.` = "Advocate",
                       `I perform Utilization Management (UM) functions at an agency serving individuals with IDD or an agency that manages services for individuals with IDD.` = "Utilization Management",
                       `I am a community living support worker for individuals with IDD.` = "Community Living Supports",
                       `I am a Habilitation Supports Waiver Director.` = "HSW Manager")) %>%
  select(-CollectorID, -StartDate, -EndDate, -starts_with("Please.rate.the."))

# Make a key to allow readable names for question content
column_key <-
  sis_survey %>%
  colnames() %>%
  as.data.frame() 
names(column_key)[1] <- "col_name"
column_key %<>%
  mutate(long_desc = recode(col_name,
                            `role` = "Respondent involvement in IDD services in Michigan",
                            `use_access` = "Make sure people receive the services they need",
                            `use_networkdev` = "Inform development of provider network",
                            `use_networkadeq` = "Evaluate provider network adequacy",
                            `use_geography` = "Identify locations to make services available",
                            `use_popdiff` = "Understand different needs across population",
                            `use_iposneeds` = "Recommend individual needs to inform IPOS",
                            `use_iposgoals` = "Identify potential goals and objectives for IPOS",
                            `use_refer` = "Identify potential service referrals",
                            `use_safetyplan` = "Help guide safety planning",
                            `use_predictcost` = "Predict costs",
                            `use_newservice` = "Forecast need for services not yet available",
                            `use_naturalsupport` = "Identify impact of losing natural supports",
                            `use_consistentauth` = "Identify services inconsistent with needs",
                            `use_itemintent` = "Align interpretation of items by assessors",
                            `use_timely` = "Ensure timely completion of assessments",
                            `use_valid` = "Review validity and reliability of assessments",
                            `sc_goals` = "Identify potential goals and objectives",
                            `sc_services` = "Identify potential service referrals ",
                            `sc_pcp` = "Prepare for or facilitate person centered planning meetings",
                            `sc_amtscpdur` = "Inform the amount, scope, and duration of services requested in the Individual Plan of Service",
                            `participant_purpose` = "The purpose of the assessment was explained clearly",
                            `participant_answers` = "The assessor was friendly and helpful in answering questions",
                            `participant_respect` = "They were treated with dignity and respect throughout the process",
                            `participant_understand` = "The assessment helped to better understand the needs of the individual being assessed",
                            `participant_results` = "They understood the results of the assessment",
                            `participant_received` = ""),
         short_desc = recode(col_name,
                            `role` = "Respondent role",
                            `use_access` = "Access to Services",
                            `use_networkdev` = "Provider Development",
                            `use_networkadeq` = "Network Adequacy",
                            `use_geography` = "Service Locations",
                            `use_popdiff` = "Population Needs",
                            `use_iposneeds` = "IPOS Needs",
                            `use_iposgoals` = "IPOS Goals",
                            `use_refer` = "Service Referrals",
                            `use_safetyplan` = "Safety Planning",
                            `use_predictcost` = "Predict Costs",
                            `use_newservice` = "Forecast Service Need",
                            `use_naturalsupport` = "Natural Support Impact",
                            `use_consistentauth` = "Consistent Authorization",
                            `use_itemintent` = "Item Intent",
                            `use_timely` = "Assessment Timeliness",
                            `use_valid` = "Validity and Reliability",
                            `sc_goals` = "Identify Goals",
                            `sc_services` = "Identify Referrals",
                            `sc_pcp` = "Prepare for PCP",
                            `sc_amtscpdur` = "Define Services",
                            `participant_purpose` = "Clear Purpose",
                            `participant_answers` = "Helpful Assessor",
                            `participant_respect` = "Treated Respectfully",
                            `participant_understand` = "Better Understanding",
                            `participant_results` = "Clear Results",
                            `participant_received` = "Received Results"))
         
use_groups <- read.csv("use_groups.csv")

```

# Summary

Feedback from stakeholders across the system
Endorses use of SIS assessment data 
List types of use
Issues that impact implementation
Caveats and inappropriate uses
Current state of implementation (Promise still in the process of being kept)

---

# What is this report about?

The Supports Intensity Scale (SIS®) is a standardized assessment tool required by the Michigan Department of Health and Human Services (MDHHS) for all Medicaid-eligible persons 18 and older who have intellectual or developmental disabilities (IDD) and who are currently receiving case management or supports coordination or respite-only services through the public behavioral health system. Use of the tool began in July of 2014, with the intent that all eligible individuals will receive an initial SIS assessment by September 30th, 2017.

In anticipation of the completion of this initial implementation phase, MDHHS sought feedback from people receiving services and supports, their family members, advocates, Prepaid Inpatient Health Plan (PIHP) employees, Community Mental Health (CMH) agencies and service providers regarding individuals’ experience with the SIS® assessment as well as current and potential uses of the data from the tool.

This document is intended to accomplish two objectives: 

* Provide a summary of feedback from individuals who participated in the survey and/or focus groups.
* Provide guidance regarding potential uses of the data from the SIS® assessment tool by stakeholders in multiple roles throughout Michigan's behavioral health system.

---

# Contributors

The guidance and thinking presented here is the result of substantial input from across the behavioral health system.  Contributions were made in the following ways...

## Focus Groups

A series of focus groups were scheduled throughout the state for individuals with intellectual and/or developmental disabilities, their family members, and their Supports Coordinators/Case Managers. The purpose of these focus groups was to understand people's current use of data from the SIS® and to understand how information gathered during the assessment might be used to inform the Person-Centered Planning process, ensuring that individuals receive supports and services according to their needs. Four focus groups were offered at locations throughout the state (*in Detroit, Petoskey, Kalamazoo, and Mount Pleasant*) to reach a broad audience while minimizing travel time for attendees. [Focus group feedback](#focuscomments) was captured in notes taken during the groups and subsequently grouped into related themes to inform this paper.

Participants in the Winter 2017 Self Determination Conference were also engaged in a focus group session and their responses are included here as well.

A total of **x** people attended these focus groups ().  While people from multiple roles and areas across the system were in attendance, there was unfortunately a relatively low turn-out for individuals served and their family members. This is especially unfortunate, since the focus groups were intended to be the primary way of gathering personal feedback from persons served and their families, in lieu of the more formal survey.

## Survey

[A survey](https://www.surveymonkey.com/r/Preview/?sm=O_2BNG1KDfFzgrTWySDftS7l01lFo5CEw11Po46B_2FKYGej_2FudfAYH_2BPr0I3SK4VNAg) was developed using *Survey Monkey* and sent by MDHHS staff via e-mail to each of the following groups on February 1, 2017:

* Advocates and Advocacy Organizations
* Habilitation Support Waiver Leads
* SIS® Assessors and Quality Leads
* CEOs of CMHSP agencies (*for distribution to IDD staff*)

A reminder e-mail was distributed on March 14, 2017 in advance of the closing of the survey on March 17, 2017.  

The survey gathered feedback from users in various roles regarding their  experience with the tool and its implementation in Michigan as well as their current and potential future uses of the SIS® assessment data.^[The list of potential uses of the data which were included as response options in the survey questions were originally identified by practitioners in the field and subsequently reviewed by MDHHS].  

`r n_distinct(sis_survey$RespondentID)` people responded to the survey from `r n_distinct(sis_survey$IP.Address)` locations across the state of Michigan. When asked which role best described their involvement in services for individuals with intellectual and developmental disabilities in Michigan, the respondents identified as follows:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
sis_survey %>%
  group_by(role) %>%
  summarize(n = n()) %>%
  mutate(role = fct_reorder(role, n, .desc = T)) %>%
  plot_ly(x = ~role, y = ~n,
          hoverinfo = 'text',
          text = ~paste(n," respondents identified their role as <br>",
                        role)) %>% 
  add_bars() %>% 
  layout(margin = list(b = 150),
         xaxis = list(title = "", tickangle = 45),
         yaxis = list(title = "Respondents in this role"))
```

By far, the largest role represented in the responses is Supports Coordinator (n = `r n_distinct(sis_survey$RespondentID[sis_survey$role == "Supports Coordinator"])`), followed by directors of programs for IDD (n = `r n_distinct(sis_survey$RespondentID[sis_survey$role == "Program Director"])`) and SIS assessors (n = `r n_distinct(sis_survey$RespondentID[sis_survey$role == "SIS Assessor or Quality Lead"])`).  All other roles had fewer than 10 responses.  

Since Supports Coordinators are a crucial part of helping to translate an individual's assessed needs into a dialogue to inform person-centered planning and eventually authorizations for services, it is encouraging to have such a strong response in this area. 

## MDHHS Review

The contents of this report have also been reviewed by key staff members within MDHHS, including:

* Amy Peckinpaugh, *Department Analyst*
* Other
* Members 
* Here

---

# Uses of the Data

## Overview

### Most common current uses

The chart below shows the number of respondents who indicated that they were already using the SIS assessment results to inform their work in a given area (i.e. they responded *"Currently using in my role"* to the specific item). See [here](#surveyitems) for a more complete description of the survey items.^[The labels in the chart shown here are summaries of the text of the question from the survey.  To see a more detailed explanation of each, you can hover above the chart or look up the related explanation in the appendix here.]

```{r, echo=FALSE, message=FALSE, warning=FALSE}

using <-
  sis_survey %>%
  select(role,starts_with("use_")) %>%
  mutate_at(.cols = vars(starts_with("use_")),
            .funs = funs(. == "Currently using in my role")) %>%
  gather(use, YorN, -role)

usable <-
  sis_survey %>%
  select(RespondentID,role,starts_with("use_")) %>%
  mutate_at(
    .cols = vars(starts_with("use_")),
    .funs = funs(current = . %in% c("Currently using in my role"),
                 could = . %in% c("Could be useful in my role"))) %>%
  select(
    RespondentID,
    role,
    ends_with("_current"), 
    ends_with("_could")
  ) %>%
  gather(use, YorN, -role, -RespondentID) %>%
  mutate(
    when = ifelse(grepl("_current", use),"Current","Potential"),
    use = gsub("_[^_]*$", "", use)
  )

using %>%
  group_by(use) %>%
  summarize(n = sum(YorN)) %>%
  ungroup() %>%
  left_join(column_key, by = c("use" = "col_name")) %>%
  droplevels() %>% ungroup() %>%
  mutate(short_desc = fct_reorder(short_desc, n, .desc = T)) %>%
  plot_ly(x = ~short_desc, y = ~n,
          hoverinfo = 'text',
          text = ~paste(n," respondents report using the data to <br>",
                        long_desc)) %>% 
  add_bars() %>% 
  layout(margin = list(b = 150),
         xaxis = list(title = "", tickangle = 45),
         yaxis = list(title = "Respondents currently using"))

```

Responses indicate that the most common use of SIS® data is to inform the individual plan of service (IPOS) and related tasks.  This is not surprising, given the large number of responses from Supports Coordinators.

### Current uses of data by role

One of the guiding principles of the questions on the survey was that uses of the data should be defined by role.  For instance, while it would be inappropriate for a supports coordinator to use solely SIS data to determine medical necessity for an individual's services, it may be entirely appropriate for a provider network manager to use the data to explore whether there are enough services in a given area to meet the needs of the entire population served.  The *role* which is using the data, and the *use* to which it is put, matters.

The visual below provides a summary of how survey respondents in specific roles said they were using the data.  In the chart, a darker blue means that a higher percent of respondents indicated they were currently using SIS data for a particular purpose.^[While the survey clearly asked respondents "How could the data from the SIS® assessment *help you to perform your role*?", it is not clear that individuals consistently responded with this guidance in mind.  Thus, a number of respondents indicated using the assessment for a purpose which clearly falls outside the scope of their role.  Furthermore, the feedback from focus groups was not captured as structured data in a way that the specific type of use to be tied to a specific role. For this reason we have not structured the subsequent sections by role.]

```{r, echo=FALSE, message=FALSE, warning=FALSE}

tst <- 
using %>%
  group_by(role, use) %>%
  summarize(n = sum(YorN),
            total = n()) %>%
  mutate(pct = n / total) %>%
  ungroup() %>%
  left_join(column_key, by = c("use" = "col_name")) %>%
  select(short_desc,role,pct) %>%
  droplevels() %>%
  group_by(short_desc) %>%
  spread(role,pct) %>%
  ungroup()

tst %>%
  select(-short_desc) %>%
  mutate_all(.funs = funs(round(. * 100, digits = 1))) %>%
  as.matrix() %>%
  plot_ly(x = colnames(.), 
          y = tst$short_desc, 
          z = ., 
          type = "heatmap",
          colors = "Blues") %>% 
  layout(margin = list(l = 170, b = 150),
         xaxis = list(title = "Role", tickangle = 45),
         yaxis = list(title = "Type of Use"))

```



**If people are not using the data, why not?**

```{r}

```

Knowing how respondents are already using the data is helpful, yet it may reflect only a few of the full range of potential uses for this data. The chart below shows the proportion of respondents who indicated that a particular use of the data would be relevant for their role.

```{r}

```

Indeed, it is likely that many potential uses were not identified in the survey responses.  As with any resource, people will use data for ingenious and unpredictable purposes.  For this reason, it is important that there be principles to guide the use of this data across the system.  The specific uses outlined in the next several sections identify promising areas for use of this information across various

---

## Use in Person-Centered Planning

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Subset use responses
use_pcp <-
  usable %>%
  left_join(column_key, by = c("use" = "col_name")) %>%
  left_join(use_groups, by = "short_desc") %>%
  filter(use_group == "Person-Centered Planning")

```

### Specific Types of Use

Since person-centered planning (PCP) is central to the entire system of services and supports, it has multiple steps and participants involved.  Thus, there are a number of distinct uses for assessment data within the person-centered planning process.

Assessment data might be used in the following ways to support person-centered planning:

* Recommend individual support needs to inform an individualized plan of service (IPOS)
* Identify potential goals and objectives for inclusion in the IPOS
* Identify potential referrals for additional assessments or relevant services
* Help guide safety planning by identifying areas of risk
* Design an individualized menu of services to support an individual to live independently in the community
* Identify individual strengths for inclusion in the IPOS
* Provide guidance regarding supports which may provide alternatives to guardianship
* Identify areas of specific personal interest or areas on which the individual wishes to focus

### Current and Potential Usage

The chart below shows the number of respondents who indicated that the data from the assessment could be relevant for them in their current roles (i.e. they responded either *"Currently using in my role"* or *"Could be useful in my role"* to the specific item).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# using a stacked bar chart with one bar for specific use and stacked by role, intended or potential uses by role

use_pcp %>%
  group_by(short_desc,long_desc,when) %>%
  summarize(n = sum(YorN),
            responders = n_distinct(RespondentID)) %>%
  ungroup() %>%
  mutate(
    pct = n / responders * 100,
    short_desc = fct_reorder(short_desc, n, .desc = T)
  ) %>%
  plot_ly(
    x = ~short_desc, 
    y = ~pct, 
    color = ~when,
    hoverinfo = 'text',
    text = ~paste(n, " out of ", responders," respondents (",
                  round(pct, digits = 1),"%)<br>",
                  "identified",ifelse(when == "Current","current","potential"),
                  " use of the data to<br>",long_desc)
  ) %>% 
  add_bars(colors = c("#132B43", "#56B1F7")) %>% 
  layout(
    margin = list(b = 150),
    barmode = 'stack',
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(title = "% indicating relevance", range = c(0, 100))
  )

```

The majority of survey respondents endorsed the SIS assessment as having usefulness as a part of the PCP process (*either current or potential*) across each of the areas indicated, from identifying needs and related goals to recommending potential services for supporting those goals.

The chart above indicates that there is already some use of this assessment data to support the PCP process, but that this is not yet widespread even among respondents who indicated that the data could be useful in this regard.  These findings from the survey were echoed in the focus group dialogues, where responses ranged from *"SIS doesn’t come up in person-centered planning."* to *"I can readily access it... review it [and] use it in the PCP process"*.

The chart below shows responses specifically from supports coordinators regarding their use of the SIS assessment in the PCP process.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

sc_uses <-
  sis_survey %>%
  select(starts_with("sc_"),-ends_with("whynot")) %>%
  gather(use,freq) %>%
  # Remove respondents not answering (d/t skip logic)
  filter(is.na(freq) == F) %>%
  left_join(column_key, by = c("use" = "col_name")) %>%
  group_by(short_desc,long_desc,freq) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(short_desc) %>%
  mutate(
    responders = sum(n),
    pct = round(n / sum(n) * 100, digits = 1),
    freq = fct_relevel(freq,"Consistently","Frequently","Occasionally","Never")
  )

sc_uses %>%
  plot_ly(
    x = ~short_desc, 
    y = ~pct, 
    color = ~freq,
    hoverinfo = 'text',
    text = ~paste(n, " out of ", responders," supports coordinators (",pct,"%)<br>",
                  "said they ", tolower(freq), " used assessment data to<br>",
                  tolower(long_desc))
  ) %>% 
  add_bars(colors = c("#83AF9B","#C8C8A9","#FC9D9A","#FE4365")) %>% 
  layout(
    margin = list(b = 50),
    barmode = 'stack',
    xaxis = list(title = ""),
    yaxis = list(title = "% indicating current use", range = c(0, 100))
  )

```

Here we see that greater than 60% of supports coordinators responding to the survey use the SIS either *Never* or only *Occasionally* for any of the activities indicated here.  When asked why they did not use it more frequently, these respondents articulated a number of issues related to the usefulness of the tool and its integration into the existing PCP process which should prompt a thoughtful reconsideration of current practice.

If one thinks of those who already using the assessment in the PCP process (i.e. those responding *Current*) as early adopters, the question for the broader system becomes how to spread the learning of that group to the rest of those who see the potential relevance of the tool (i.e. those responding *Potential*) and ultimately to demonstrate the usefulness of the assessment for those who did not originally see its potential (i.e. those who did not endorse either *Current* or *Potential* use).^[This is a common challenge with the implementation and   diffusion of any change, which is impacted by the nature of the innovation itself, communication related to the change, time, and a social system which supports the use of the innovation.]

### Considerations

A number of helpful considerations were identified by focus group attendees during their dialogues.  These themes are summarized below.

#### Integration into Process

While the information from the assessment may have potential, making it useful in reality will require a purposeful integration of the information into the PCP process for individuals receiving services.  

An integrated process would be one that:

* _Summarizes relevant results_: Distills information from the assessment which is relevant to the individual person, rather than merely providing an abstract score (too little information) a list of all question responses (too much information).
* _Makes results available_: The information should be easily available to those involved in the PCP process in a timely manner.
* _Honors individual choice_: The use of the information from the assessment should provide a meaningful set of options without automating the decision-making process or taking away individual control.

Feedback from focus groups consistently stressed that this integration was crucial, but was far from being the standard of practice in current person-centered planning.  The excerpted feedback below attests to this:

> *"It has aspects that can improve the PCP process. On the state level, we [as advocates] wanted to focus more on improving PCP process. Instead of just saying we have in law that PCP is how we drive services, if there could be a merging [of assessment data into a more consistent PCP process] that could be a beautiful thing."*

> *"SIS doesn’t come up in person-centered planning. We’re not told as supports coordinators that we need to integrate it into the process... Only when you start talking about authorizations does the SIS come up."*

> *"We sat down for 90 minutes with a CM to make that connection. She’s the only one... that uses the SIS and aligns it with PCP. She’s tried to show new CM’s, and they’re like... well there’s just a lot that they have to do."*

For this integration to become a reality, several related logistical issues will need to be resolved:

1. Consistent Integration into PCP Process
2. Consistent Integration into EMR Workflow supporting PCP Process
3. Scheduling of Assessments to Align with PCP
4. Consistent Training of Facilitators
5. Guidance Regarding Use of Data in 2nd and 3rd Year after Initial Assessment

#### Benefits of Use

A number of distinct benefits were identified related to the use of assessment data in the person-centered planning process:  

* Insight into individual needs and strengths for all treatment team members
* The assessment was felt to be well aligned with the content of other assessments, and to provide support for these
* The consistent identification of health and safety needs 
* In addition to relatively stable needs, the assessment also allows for a focus on more dynamic needs which may change over time (e.g. behavioral needs)
* Overcoming preconceptions of family members

Feedback from respondents included the following:

> *"It took 7 times of hearing my client say the same thing, the assessor acknowledged it, and then it clicked for me that it was a necessary goal. [I had spent] 13 years of working with that person without having that realization."*

> *"We’re having the SIS and all this stuff is coming out about what she can do. By the time we have the PCP, the sister and consumer was able to come in and realize... if she has aspirations to do her own thing, she’s going towards that."*

> *"A person who lives with their parent currently, being able to say 'It’s not that I can’t cook, it’s that you don’t allow me to do things.' ...Not what’s happening right now, but what could happen."*

#### Caveats for Use

In addition to the benefits noted above, several important cautions were communicated by respondents:

* _Personal priorities change_: While the SIS assessment allows a person to tag specific support areas as *"important to"* them, focus group participants cautioned that these items should not be interpreted as accurate indicators of a person's current priorities, since these will naturally change with time.
* _Mismatch between needs and preference_: While the SIS assessment identifies both support needs and individual preferences (*in the form of "important to" items tagged in the assessment*), the relationship between these is not always straightforward to interpret.  For instance, a person may "need" a high level of supports in employment, but this would not be relevant for individual planning if the person was not interested in pursuing employment.  On the other hand, essential safety-related items (*eating, for example*) may need to be incorporated into the person-centered planning process even if they are not explicitly endorsed by the person.

---

## Use in Managing a Provider Network

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Subset use responses
use_prov <-
  usable %>%
  left_join(column_key, by = c("use" = "col_name")) %>%
  left_join(use_groups, by = "short_desc") %>%
  filter(use_group == "Managing a Provider Network")

```

### Specific Types of Use

Tailoring services to the needs of a specific individual is a hallmark of person-centered planning, but there is an additional level of complexity required to support the design of provider networks to meet the precise needs of the populations they serve.  Assessment data from the SIS® has been identified as having potential applications in the following areas related to provider network management:

* Inform the development of provider network service offerings. 
* Evaluate provider network adequacy relative to the specific needs of the population in a given area.         
* Identify locations to make services available for individuals within an accessible distance from their homes.
* Forecast needs for services not yet available. 
* Provide consistent guidance to service providers related to the scope, duration and intensity of services authorized by the IPOS.
* Assist providers to develop an enhanced array of services to support compliance with the HCBS Final Rule.

### Current and Potential Usage

The chart below shows the number of survey respondents who indicated that the data from the assessment could be relevant for them in their current roles (i.e. they responded either *"Currently using in my role"* or *"Could be useful in my role"* to the specific item).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# using a stacked bar chart with one bar for specific use and stacked by role, intended or potential uses by role

use_prov %>%
  group_by(short_desc,long_desc,when) %>%
  summarize(n = sum(YorN),
            responders = n_distinct(RespondentID)) %>%
  ungroup() %>%
  mutate(
    pct = n / responders * 100,
    short_desc = fct_reorder(short_desc, n, .desc = T)
  ) %>%
  plot_ly(
    x = ~short_desc, 
    y = ~pct, 
    color = ~when,
    hoverinfo = 'text',
    text = ~paste(n, " out of ", responders," respondents (",
                  round(pct, digits = 1),"%)<br>",
                  "identified",ifelse(when == "Current","current","potential"),
                  " use of the data to<br>",long_desc)
  ) %>% 
  add_bars(colors = c("#132B43", "#56B1F7")) %>% 
  layout(
    margin = list(b = 150),
    barmode = 'stack',
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(title = "% indicating relevance", range = c(0, 100))
  )

```

While the majority of survey respondents endorsed the SIS assessment as having usefulness in managing and developing provider networks (*either current or potential*) across each of the areas indicated, a higher proportion marked these as *potential* uses, indicating that these applications of the data are not yet developed.

This was echoed in focus group dialogues, such as the provider network manager who noted that:

> *"This could be useful when we assess... is our paneled provider [network] sufficient to meet the needs in this area?"*

### Considerations

A number of helpful considerations were identified by focus group attendees during their dialogues.  These themes are summarized below.

#### Use in Service Coordination and Transitions

The usefulness of a standardized assessment such as the SIS to serve as a common language for coordination and transtions across providers and settings was raised consistently by focus group participants, who identified several distinct types of applications:

* Transitions between settings and services
* Coordination of concurrent services across multiple providers
* Transitions between supports coordinators (*e.g. due to employee turnover*)

The statements below reinforce a growing body of [research](https://scholar.google.com/scholar?q=community+care+transitions) and [effort](https://www.medicaid.gov/medicaid/quality-of-care/improvement-initiatives/care-transitions/index.html) surrounding coordination and transitions of care:

> *"They used to do SIS for all new intakes, and it was extremely helpful for supports coordinators..."*

> *"For new supports coordinators, the SIS would be helpful. When I’ve known them over 10 years, I can probably help the assessor out."*

> *"If a new case manager comes on, and they have to understand their entire caseload they’ve been tasked with, they can start to understand them better with access to thorough assessments"*

#### Messaging to People Receiving Services

A number of participants noted that their personal messaging to clients regarding the usefulness of the SIS assessment included discussions about improving the relevance of services.  One participant noted that:

> *"I framed [the SIS assessment to clients as a tool that] helps us to be able to provide the services you need in the future."*

At the current time, however, the provider network has not yet evolved to incorporate this information, and this can lead to frustration at the individual level:

> *"People ask,* How might that help me? *with questions related to job services, for example, when there are no services available"*

It will be important to continue to share ways in which individual needs and interests are being considered to shape a more responsive array of services.

---

## Use in Assuring Access and Equity

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Subset use responses
use_um <-
  usable %>%
  left_join(column_key, by = c("use" = "col_name")) %>%
  left_join(use_groups, by = "short_desc") %>%
  filter(use_group == "Assuring Access and Equity")

```

### Specific Types of Use

Of all the potential uses of data from the SIS® assessment, this was one of the most commonly endorsed during focus group forums.  Assessment data from the SIS® has been identified as having potential applications in the following areas related to assuring access to services and ensuring equity of access for all persons receiving services:

* Ensure people receive the services they need 
* Identify services inconsistent with assessed needs
* Inform the prioritization of access to available waiver slots

### Current and Potential Usage

The chart below shows the number of respondents who indicated that the data from the assessment could be relevant for them in their current roles (i.e. they responded either *"Currently using in my role"* or *"Could be useful in my role"* to the specific item).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# using a stacked bar chart with one bar for specific use and stacked by role, intended or potential uses by role

use_um %>%
  group_by(short_desc,long_desc,when) %>%
  summarize(n = sum(YorN),
            responders = n_distinct(RespondentID)) %>%
  ungroup() %>%
  mutate(
    pct = n / responders * 100,
    short_desc = fct_reorder(short_desc, n, .desc = T)
  ) %>%
  plot_ly(
    x = ~short_desc, 
    y = ~pct, 
    color = ~when,
    hoverinfo = 'text',
    text = ~paste(n, " out of ", responders," respondents (",
                  round(pct, digits = 1),"%)<br>",
                  "identified",ifelse(when == "Current","current","potential"),
                  " use of the data to<br>",long_desc)
  ) %>% 
  add_bars(colors = c("#132B43", "#56B1F7")) %>% 
  layout(
    margin = list(b = 150),
    barmode = 'stack',
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(title = "% indicating relevance", range = c(0, 100))
  )

```

The majority of survey respondents endorsed the SIS assessment as having usefulness in assuring access and equity (*either current or potential*) across each of the areas indicated.  A higher proportion of respondents indicated that they currently used the assessment to ensure access to services (i.e. *ensure people receive the services they need*) than to identify unnecessary services (i.e. *identify services inconsistent with assessed needs*).

### Considerations

A number of helpful considerations were identified by focus group attendees during their dialogues.  These themes are summarized below.

#### A Difficult Balance

While stakeholders consistently endorsed the goal of improving access to services, there was an acknowledged complexity in ensuring that access was distributed equally within budgetary constraints.

Broadly speaking, there were four different types of decisions that participants alluded to in relation to service access:

* Renew access (*Continue access to existing service/benefit*)
* Identify need (*Facilitate access to new service/benefit*)
* Deny request (*Prevent access to new service/benefit*)
* Remove access (*Remove access to existing service/benefit*)

Of these, the latter two types of decisions were of greater concern to participants, whose primary concern is with individual people and not with budgets.  However, participants also acknowledged the tension that a disposition toward the first two decision types will tend to use up available resources more quickly.

Respondents from various roles acknowledged this tension when describing their current uses and future hopes for information from the tool to assure access and equity:

> *"I know our SIS assessor always makes it a point to say, this does not decrease services, but could add extra emphasis as to why they do or don’t need a service."*

> *"[We] have had people bring up the SIS score when it doesn’t jive with the level of services being requested."*

> *"We use [the SIS results] as a talking point to talk about prioritizing our Hab Support Waiver spots. It isn’t an end-all, but helps us to prioritize."*

> *"As new people come in, we have a chance to give equality of service for similar level of needs. And wherever they might go, they’ll be able to see that they have the tool completed and understand that level of need."*

> *"When I know that the SIS is going to be compiled and go to money-makers... there’s people at the local level that can allocate resources in a more independent way..."*

---

## Use in Understanding Population Needs

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Subset use responses
use_pop <-
  usable %>%
  left_join(column_key, by = c("use" = "col_name")) %>%
  left_join(use_groups, by = "short_desc") %>%
  filter(use_group == "Understanding Population Needs")

```

### Specific Types of Use

In distinction from uses of the data discussed above, such as access and person-centered planning, this section pertains to uses of data which are especially relevant at the state and PIHP level.  Assessment data from the SIS® has been identified as having potential applications in the following areas related to population needs:

* Understand different needs across population
* Predict costs                               
* Identify impact of losing natural supports due to an aging population of parents
* Implement federal and other requirements in a way which optimizes their benefit for the entire population of individuals receiving services

### Current and Potential Usage

The chart below shows the number of respondents who indicated that the data from the assessment could be relevant for them in their current roles (i.e. they responded either *"Currently using in my role"* or *"Could be useful in my role"* to the specific item).

```{r, echo=FALSE, message=FALSE, warning=FALSE}

use_pop %>%
  group_by(short_desc,long_desc,when) %>%
  summarize(n = sum(YorN),
            responders = n_distinct(RespondentID)) %>%
  ungroup() %>%
  mutate(
    pct = n / responders * 100,
    short_desc = fct_reorder(short_desc, n, .desc = T)
  ) %>%
  plot_ly(
    x = ~short_desc, 
    y = ~pct, 
    color = ~when,
    hoverinfo = 'text',
    text = ~paste(n, " out of ", responders," respondents (",
                  round(pct, digits = 1),"%)<br>",
                  "identified",ifelse(when == "Current","current","potential"),
                  " use of the data to<br>",long_desc)
  ) %>% 
  add_bars(colors = c("#132B43", "#56B1F7")) %>% 
  layout(
    margin = list(b = 150),
    barmode = 'stack',
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(title = "% indicating relevance", range = c(0, 100))
  )

```

### Considerations

## Use in Advocacy

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Subset use responses
use_advoc <-
  usable %>%
  filter(role == "Advocate") %>%
  left_join(column_key, by = c("use" = "col_name")) %>%
  left_join(use_groups, by = "short_desc") 

```

### Specific Types of Use

Advocates perform a role with a broad scope.  Since their interest focuses on the individual person's quality of life, they might become interested in any of the uses of the assessment mentioned above, but approach those uses from the specific perspective of their impact on the individual. 

* Provide guidance regarding supports which may provide alternatives to guardianship
* Use in fair hearing


### Current and Potential Usage

The chart below shows the number of advocates who indicated that the data from the assessment could be relevant for them in their current roles (i.e. they responded either *"Currently using in my role"* or *"Could be useful in my role"* to the specific item).

```{r, echo=FALSE, message=FALSE, warning=FALSE}

use_advoc %>%
  group_by(use_group,short_desc,long_desc) %>%
  summarize(n = sum(YorN),
            responders = n_distinct(RespondentID)) %>%
  ungroup() %>%
  mutate(
    pct = n / responders * 100,
    short_desc = fct_reorder(short_desc, n, .desc = T)
  ) %>%
  plot_ly(
    x = ~short_desc, 
    y = ~pct, 
    color = ~use_group,
    hoverinfo = 'text',
    text = ~paste(n, " out of ", responders," advocates identified<br>",
                  "current or potential use of the data to<br>",long_desc)
  ) %>% 
  add_bars() %>% 
  layout(
    margin = list(b = 150),
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(title = "% indicating relevance", range = c(0, 100))
  )

```

### Considerations

> *"We advocate for individuals with I/DD, [and] while we don’t typically sit in on SIS, it helps us to articulate and understand their situation"*

> *"The SIS... was brought up in a Fair Hearing, and it allowed to increase services. That’s an advocacy tool use. Went all the way up the chain with it."*

## Inappropriate Uses

# What issues impact use of the data?

While data is stored in databases and defined in specifications, the *meaning* of the data is influenced by any number of issues which include perceptions of those collecting, analyzing and disseminating the data, and how the data is explained. 

## Messaging related to use of the SIS

It is worth noting that, during the focus groups, respondents did not draw a clear line between the official messaging provided by MDHHS and the informal messaging communicated within their own organizations or teams.  The issues with messaging noted here reflect individual perceptions rather than a comprehensive review of formal guidance to date.^[For instance, people acknowledged that multiple resources had been made available during the initial roll-out of the assessment.  One respondent noted that: *"The Department had trainers around the area, so there were presentations...there were quite a few of those"*  Nevertheless, a number of individuals were not aware of these resources and therefore noted a lack of clarity in messaging.]

### Lack of clarity around how data will be used

### Fear of use to decrease access to services {#fearlimit}

### Mandate

Lack of clarity in how to integrate into PCP process during initial rollout
Lack of clarity around requirement of SIS
Miscellaneous
Communication with families could be improved
Informing Service Availability
MDHHS training available
Perception of assessment as indicating lack of trust in supports coordinators
Usefulness in service planning
Process for requesting an updated assessment
Usefulness to develop goals
Recognition of limits
Equity


## Facilitation of SIS assessment

One of the primary experiences which impacts the implementation and interpretation of the SIS assessment data is the experience which they have had with facilitation of the assessment, or what has been communicated to them by others about this experience.^[Please note that the focus of this paper treats the process of the assessment only to the extent that it impacts the interpretation of the resulting dataset or the willingness of the broader system to use the data.  This is not an evaluation of the content of training about how the tool should be implemented, but a summary of feedback from stakeholders about how they have experienced it.]

`r length(sis_survey$RespondentID[sis_survey$participated == "Yes and I remember the experience"])` survey respondents indicated that they had participated in a SIS assessment and remembered their experience.  These individuals were asked to respond to several questions regarding that experience, and their responses are summarized in the chart below:

```{r, echo=FALSE, message=FALSE, warning=FALSE}

participant <-
  sis_survey %>%
  select(starts_with("participant_"),-ends_with("received")) %>%
  gather(area,response) %>%
  # Remove respondents not answering (d/t skip logic)
  filter(is.na(response) == F) %>%
  left_join(column_key, by = c("area" = "col_name")) %>%
  group_by(short_desc,long_desc,response) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  droplevels() %>%
  group_by(short_desc) %>%
  mutate(
    responders = sum(n),
    pct = round(n / sum(n) * 100, digits = 1),
    response = fct_relevel(response,"Strongly Agree","Agree","Disagree","Strongly Disagree")
  )

participant %>%
  plot_ly(
    x = ~short_desc, 
    y = ~pct, 
    color = ~response,
    hoverinfo = 'text',
    text = ~paste0(n, " out of ", responders," respondents (",pct,"%)<br>",
                  "who attended an assessment ", tolower(response), "d that<br>",
                  tolower(long_desc))
  ) %>% 
  add_bars(colors = c("#83AF9B","#C8C8A9","#FC9D9A","#FE4365")) %>% 
  layout(
    margin = list(b = 80),
    barmode = 'stack',
    xaxis = list(title = ""),
    yaxis = list(title = "% endorsing", range = c(0, 100))
  )

```

Overall, people's experience of the facilitation of the SIS assessment appeared to be quite positive.  The two areas which had less favorable responses had to do with the clarity of the assessment results and the contribution of the assessment process to their overall understanding of the individual being assessed.

That being said, it is worth noting that the majority of respondents who had participated in a SIS assessment indicated that the results of the assessment were clear and that they had a better understanding of the person being assessed as a result.

## Format of the SIS tool

## Completeness of data

# Supplement A: Survey Items {#surveyitems}

The table below shows more detailed descriptions of the survey items which are referenced in the charts throughout this document.  These descriptions can also be seen by hovering over the charts in the web version of the report.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(DT)

column_key %>%
  filter(grepl("^use_", col_name)) %>%
  select(short_desc, long_desc) %>%
  datatable(
    caption = 'Shortened Descriptions of Assessment Uses',
    colnames = c('Short Name', 'Survey Response Detail'),
    rownames = FALSE,
    options = list(dom = 't',pageLength = 16)
  )

```

# Supplement B: Focus Group Comments {#focuscomments}

The searchable table below includes all comments captured in notes during the focus group sessions.  These have been grouped together and classified by themes and by the type of usage they relate to.  In certain cases, quotes are paraphrased based on what could be captured by the notetaker during the session.  Some comments may not have been captured due to the pace of the conversation during portions of the focus groups.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(DT)

sis_focus %>%
  mutate(
    theme = as.factor(theme),
    subtheme = as.factor(subtheme),
    quote = str_replace_all(quote, "[[:punct:]]", " ")
  ) %>%
  select(theme,subtheme,quote,use) %>%
  arrange(theme,subtheme) %>%
  datatable(
    caption = 'Annotated Comments from Focus Groups',
    colnames = c('Theme','Sub-theme','Quote','Type of Use'),
    rownames = FALSE,
    filter = 'top',
    options = list(pageLength = 5,lengthMenu = c(5, 10, 20))
  )

```


                                       

